Block bootstrap consistency\\under weak assumptions

### 2016-06-10 Paper.tex
[#25] Fix duplicate equation refernce in Theorem 1
Closes #25.

Theorem~\ref{T1} presents the main result, asymptotic normality of
the distribution of bootstrapped sums. This paper adopts the
standard notation that $\E^{*}$, $\var^{*}$, etc.\ are the usual
operators with respect to the probability measure induced by the
bootstrap and will use explicit stochastic array notation
for precision.  Also note that all results are presented for the
scalar case but generalize immediately to random vectors.  All of the
proofs are presented in the appendix; only proofs for the Stationary
Bootstrap are presented, since proofs for the other methods are
similar and easier to construct.  All limits are taken as $n \to \infty$
unless otherwise noted and $\lVert \cdot \rVert_r$ denotes the
$L_r$-norm.

\begin{thm}\label{T1}
  Suppose the following conditions hold.
  \begin{enumerate}
  \item $X_{nt}$ is $L_2$-\ned\ of size $-1/2$ on an array
    $\{V_{nt}\}$ that is either strong mixing of size $-r/(r-2)$ or
    uniform mixing of size $-r/2(r-1)$, with $r > 2$.  The
    \ned\ magnitude indices are denoted $\{d_{nt}\}$.
  \item The array $\mu_{nt} - \bar \mu_n$ is uniformly bounded
    and $\sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2 \to 0$,
    where $\E X_{nt} = \mu_{nt}$ and $\bar{\mu}_n = n^{-1} \sum_{t=1}^n
    \mu_{nt}$. Moreover, $\sqrt{n} \| \bar{X}_{n} - \bar\mu_n \|_2
    \to \sigma > 0$, with $\bar X_n = (1/n) \sum_{t=1}^n X_{nt}$.
  \item There exists an array of positive real numbers $\{c_{nt}\}$
    such that $(X_{nt} - \mu_{nt})/c_{nt}$ is uniformly $L_r$-bounded
    and $c_{nt}$ and $d_{nt}/c_{nt}$ are uniformly bounded in $n$
    and~$t$.
  \item $X_{nt}^{*}$ is generated by the Stationary Bootstrap with
    geometric block lengths with success probability $p_n$, $p_n = c
    n^{-a}$ and $a,c \in (0,1)$, or by the Moving or Circular Block
    bootstrap with block length $M_n$ such that $M_n \sim n^a$ for
    $a \in (0,1)$.  Let $M_{ni}$ be the block length of the $i$th
    block, $i=1,\dots,J_n$, and define $K_{n0} = 0$ and $K_{nj} =
    \sum_{i=1}^j M_{ni}$. The last block, $M_{n,J_n}$, is defined as
    $K_{n,J_n} - K_{n,J_n-1}$, so $K_{n,J_n} = n$ a.s.
  \end{enumerate}
  Then $\hat\sigma^{*2} \to^p \sigma^2$,
  \begin{equation}\label{eq:36}
    \sup_x \big\lvert \pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \E^* \bar X_n^*) \leq x \big]
    - \pr \big[\sqrt{n}(\bar X_{n} - \E \bar X_n) \leq x \big] \big\rvert \to^p 0,
  \end{equation}
  and
  \begin{equation}
    \label{eq:2}
    \sup_x \big\lvert \pr^{*}\big[
    \sqrt{n}(\bar X_{n}^{*} -  \E^* \bar X_n^*) \big/ \hat\sigma_n^{*}
    \leq x \big] - \Phi(x) \big\rvert \to^p 0,
  \end{equation}
  where $\Phi$ is the \allcaps{cdf} of the Standard Normal
  distribution and
    \begin{equation}
      \label{eq:3}
      \hat{\sigma}_n^{*2} = \tfrac{1}{n} \sum_{j=1}^{J_n}
      \Big\{\sum_{t=K_{n,j-1}+1}^{K_{nj}} (X_{nt}^{*} - \bar X_{n}^{*})\Big\}^2.
    \end{equation}
\end{thm}

@@

Theorem~\ref{T1} relies on a general insight about the
variance of the sample mean under the bootstrap-induced
distribution. It is well-known that a key step in proving the \clt\
for arbitrary dependent processes is demonstrating that the squared
elements converge to a positive and finite limit; i.e.\ if
$\{Z_{nj};~j=1,\dots,J_n\}$ is a representative stochastic array,
$\sum_j Z_{nj}^2 \to^p \sigma^2$ is an important necessary condition
for $\sum_j Z_{nj} \to^d N(0,\sigma^2)$.  (See Section 3.2 of
\citealp{HaH:80}, for further discussion.)  For martingale difference
arrays, each $Z_{nj}$ is one of the original random variables $X_{nt}$
(typically normalized by $1/\sqrt{n}$), but for other forms of
dependence (\ned\ or mixingale arrays, for example, as in
\citealp{Jon:97}) each $Z_{nj}$ is a contiguous block of the original
random variables,
\begin{equation*}
  Z_{nj} = \tfrac{1}{\sqrt{n}} \sum_{t=(j-1) M_n +1}^{j M_n} (X_{nt} - \mu_{nt}),
\end{equation*}
that adds up to the original summation (plus potentially a negligible
residual) so $\sum_j Z_{nj}= (1/\sqrt{n}) \sum_t (X_{nt} - \mu_{nt}) +
o_p(1)$. In \citet{Jon:97}, for example, the \clt\ for mixingale
arrays assumes that there exists a sequence of blocks such that $\sum_j
Z_{nj}^2$ converges i.p., and the \ned\ \clt\ establishes conditions
under which such a $Z_{nj}$ exists.

Our insight is that the expectation of squared blocks of the bootstrap
process can be expressed as a sequence of \emph{contiguous} blocks of
the original process, so the arguments that establish convergence of
the original squared blocks can be applied with only minor changes to
the bootstrapped blocks. Consider the Moving Blocks Bootstrap,%
\footnote{To make this presentation as simple as possible, assume for
  now that $n = M_n J_n$ exactly.} %
for example, and let
\begin{equation*}
  Z_{nj}^* = \tfrac{1}{\sqrt{n}} \sum_{t=(j-1) M_n + 1}^{j M_n} (X_{nt}^* - \bar X_n).
\end{equation*}
Then, conditional on the data, the $Z_{nj}^{*2}$ are independent can
be expected to obey an \lln, so %
$\sum_j (Z_{nj}^{*2} - \E^* Z_{nj}^{*2}) \to^p 0$ %
and the CLT for the bootstrapped array requires $\sum_j \E^*
Z_{nj}^{*2}$ to converge to a positive and finite limit.  But, since
$\E^*$ only averages over the starting point of each block, we have
\begin{align*}
  \E^* Z_{nj}^{*2} &=
  \tfrac{1}{n} \sum_{\tau = 0}^{n-M_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= \tau + 1}^{\tau + M_n} (X_{nt} - \bar X_n) \Big)^2 \\
  &= \tfrac{1}{n} \sum_{\tau_0 = 0}^{M_n-1}
  \sum_{j=1}^{J_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= (j - 1) M_n + \tau_0 + 1}^{j M_n + \tau_0} (X_{nt} - \bar X_n) \Big)^2
\end{align*}
after grouping blocks separated by $M_n$ periods. For each $\tau_0$,
the summation can be expected to converge in probability through the
same arguments that were used to establish the \clt\ for the original
array.%
\footnote{Some of the details of the argument will typically need to
  change because the original \clt\ only requires convergence for
  $\tau_0 = 0$, but these details are often incidental to the original
  argument.}%
\footnote{Lemmas~\ref{L3} and
  \ref{L5} are particularly
  strong demonstrations of this argument.} %
A similar representation is available for the Circular and Stationary
bootstraps.

In short, the basic approach that we use to prove Theorem~\ref{T1}
is based on a fundamental connection between the second moments of the
bootstrap process and the sum of squared blocks of the original
array. Even though the details of our proof rely on specific
techniques for \ned\ arrays, this connection implies that block
bootstraps are typically consistent when the original dependent array
obeys the \clt\ and the connection should be useful for proving
consistency of the bootstrap under other dependence conditions.

Theorem~\ref{T1} can also be extended to give an \fclt\
using arguments from \citet{JoD:00b}. We show in
Theorem~\ref{T2} that the partial sum
of the bootstrapped process obeys an \fclt\ and can be used to derive
critical values for other test statistics under the same assumptions
as Theorem~\ref{T1}. For this result, define the
following partial sums,
\begin{align*}
W_n(\gamma) &=
\tfrac{1}{\sqrt{n}} \sum_{t=1}^{\lfloor \gamma n \rfloor}
(X_{nt} - \bar\mu_n) &&\text{and}&
W_n^{*}(\gamma) &= \tfrac{1}{\sqrt{n}}
\sum_{t=1}^{\lfloor \gamma n \rfloor}
(X^*_{nt} - \mu_n^{*}).
\end{align*}
Also let $W$ denote standard Brownian Motion and $\sigma W$ denote
Brownian Motion scaled by the constant $\sigma$.

\begin{thm}\label{T2}
Suppose that the conditions of Theorem~\ref{T1} hold
and let $\dist$ be any distance function that metricizes weak
convergence. Then
\begin{equation}\label{eq:7}
\pr^{*}[\dist(W_n^{*}, \sigma W) > \delta] \to^p 0
\end{equation}
for any positive  $\delta$.
If, in addition, $\sup_{t=1,\dots,n}|\mu_{nt} -
\bar \mu_n| = o(1/\sqrt{n})$ and
\begin{equation}\label{eq:4}
n^{-1} \sum_{s,t=1}^{\lfloor \gamma n \rfloor} \cov(X_{ns}, X_{nt}) \to \sigma^2 \gamma
\end{equation}
for all $\gamma \in [0,1]$, then
\begin{equation}\label{eq:8}
\pr[\dist(W_n, \sigma W) > \delta] \to 0
\end{equation}
for any positive $\delta$.
\end{thm}
If both~\eqref{eq:7} and~\eqref{eq:8} hold then the bootstrap can be
used to approximate the distribution of partial sums.
Note that Theorem~\ref{T2} imposes
stronger assumptions for the original process
than for the bootstrapped process. Without~\eqref{eq:4}, the
partial sum of the original observations converges to a transformed
Brownian Motion with a different covariance process. The bootstrapped
partial sum, on the other hand, always converges to standard (but
potentially scaled) Brownian Motion because the resampling strategies
ensure that the bootstrapped process is globally covariance stationary.

@@

\subsection*{Proof of Theorem~\ref{T1}}
{%
\newcommand{\Db}[1][j]{D_{n#1}^{*}}
\newcommand{\Zsum}{\sum_{j=1}^{J_n} \Zb / \sigma_n}
First we prove that
\begin{equation}\label{eq:37}
  \sup_x \big\lvert \pr^{*}\big[
  \sqrt{n}(\bar X_{n}^{*} -  \bar X_n) \big/ \sigma_n^{*}
  \leq x \big] - \Phi(x) \big\rvert \to^p 0
\end{equation}
where $\sigma_n^{*2} = n \E^* (\bar X_n^* - \bar X_n)^2$.
Rewrite $\sqrt{n}(\bar X_{n}^{*} - \bar X_n)$ as in
Equation~\eqref{eq:16}, so
\begin{equation*}
  \tfrac{1}{\sqrt{n}} \sum_{t=1}^n (X_{nt}^* - \bar X_n) =
  \sum_{j=1}^{J_n} Z_{nj}^*
\end{equation*}
and $\{\Zb / \sigma_n^*, \Gs_{nj}\}$ is a martingale difference
array. Moreover,
\begin{equation}
  \label{eq:18}
  \Pm^{*} \Big[\sum_{j=1}^{J_n} \Zb / \sigma_n^* \leq x \Big] - \Phi(x) \to^p 0,
\end{equation}
for all $x$ if $\sigma_n^{*2} \to^p \sigma^2$ (which ensures that
$\sigma_n^{*2}$ is uniformly a.s.\ positive and holds by
Lemma~\ref{L3}) and the following
two conditions hold for all positive $\epsilon$:
\begin{equation}
  \label{eq:19}
  \sum_{j=1}^{J_n} \Em^{*} \big(Z_{nj}^{*2} \ind\{Z_{nj}^{*2}  >
  \epsilon\}\big) \to^p 0
\end{equation}
and
\begin{equation}
  \label{eq:20}
  \Pm^{*}\Big[\ \Big|\sum_{j=1}^{J_n} Z_{nj}^{*2} - \sigma_n^{*2}
  \Big|\ > \epsilon \Big] \to^p 0
\end{equation}
since~\eqref{eq:19} and~\eqref{eq:20} ensure that $\Zb/\sigma_n^*$ obeys
a martingale difference \clt\ \citep[e.g.][Theorem 3.3]{HaH:80}.%
\footnote{Conditional on $X_{n1},\dots,X_{nn}$, $J_n$, and
  $M_{n1},\dots,M_{n,J_n}$, the only stochastic components of $\Zsum$
  are the start periods of each block, which are
  $\discreteuniform(1,\dots,n)$ and are independent of all of the
  other random variables in the information set used for conditioning.
  Consequently, $\Pm^*$ is a regular conditional probability and
  arguments like Hall and Heyde's (1980) Theorem 3.3
  apply without modification on this probability measure. See
  also Section 23.2 of \citet{Vaa:00}.
  Also note that Hall and Heyde's Theorem 3.3 as stated imposes an
  additional restriction on the sigma-fields. However, as Hall and
  Heyde discuss on pages 59 and 63--64, that condition is unnecessary
  here because $\sigma_n^{*2}$ is measurable with respect to
  all of the $\Gs_{nj}$.} %

For~\eqref{eq:20}, $Z_{nj}^*$ and $Z_{nk}^*$ (when $k \neq j$) are
conditionally uncorrelated given $X_{n1},\dots,X_{nn},$ and $\Ms$, which
implies
\begin{equation*}
  \sum_{j=1}^{J_n} Z_{nj}^{*2} - \sigma_n^{*2} =
  \sum_{j=1}^{J_n} \Big( Z_{nj}^{*2} - (1/J_n) \E^* \sum_{j=1}^{J_n} Z_{nj}^{*2} \Big)
\end{equation*}
almost surely. But
\begin{equation*}
   \Big\{\tfrac{n}{M_{nj}} \Big(Z_{nj}^{*2} - (1/J_n) \E^* \sum_{j=1}^{J_n} Z_{nj}^{*2}\Big),
   \ \Gs_{nj} \Big\}
\end{equation*}
is a uniformly-integrable martingale difference array
by Lemma~\ref{L5}
and satisfies the \lln\ (e.g., \citealp{Dav:94}, Theorem 19.7). So
this sum converges to zero in conditional probability, proving~\eqref{eq:20}.

To prove~\eqref{eq:19}, it suffices to show that
\begin{equation}\label{eq:21}
  \Em \sum_{j=1}^{J_n} \Em^{*}\big(Z_{nj}^{*2} \ind\{Z_{nj}^{*2} > \epsilon\}\big) \to^p 0.
\end{equation}
For any $j$ and $m$ and for large enough $n$,
\begin{align}
\Em \Em^{*}\big((Z_n(u_{nj}, m)^2n/m) &\ind\{Z_n(u_{nj}, m)^2n/m > \epsilon n/m\}\big) \notag \\
&= \E \big((Z_n^{*}(0, m)^2n/m) \ind\{Z_n^{*}(0, m)^2n/m > \epsilon n/m\}\big) \notag \\
&\leq B(\epsilon n/m)\label{eq:5}
\end{align}
where $B$ is some finite monotone function such that $B(x) \to 0$ as
$x \to \infty$; the existence of a function $B$ that satisfies these
conditions (and does not depend on $n$ or $m$) is a consequence of
Lemma~\ref{L5},
Equation~\eqref{eq:30}. Since $J_n$ and $M_{nj}$ are both $\Ms$-measureable
and independent of the $u_{nj}$'s and $X_{nt}$'s,~\eqref{eq:5} implies that
\begin{align*}
\Em &\sum_{j=1}^{J_n} \Em^{*}\big(Z_{nj}^{*2} \ind\{Z_{nj}^{*2} > \epsilon\}\big) \\
& = \sum_{j=1}^{J_n} (M_{nj}/n) \Em \Em^{*}
\big((Z_n(u_{nj}, M_{nj})^2 n/M_{nj})
\ind\{Z_n(u_{nj}, M_{nj})^2n/M_{nj} > \epsilon n/M_{nj}\}\big) \\
& \leq \sum_{j=1}^{J_n} (M_{nj}/n) \, B(\epsilon n/M_{nj}) \\
& \leq \max_{j = 1,\dots,J_n} B(\epsilon n / M_{nj}) \sum_{i=1}^{J_n} M_{ni} / n \\
& = B(\epsilon n / \max_{j = 1,\dots,J_n} M_{nj}) \\
& \to^p 0 \text{\ as\ } n \to \infty.
\end{align*}
The first equality follows from simple algerbra and the definition
of the $u_{nj}$'s. The first inequality is a
consequence of~\eqref{eq:5}.  The second equality holds by
monotonicity of $B$. And convergence in probability holds by
Lemma~\ref{L1}, completing the proof
of~\eqref{eq:19}. The Dominated Convergence Theorem and~(\ref{eq:18})
then ensure that
\begin{equation}\label{eq:22}
  \pr^{*} \Big[\sum_{j=1}^{J_n}
  Z_{nj}^{*} / \sigma_n^* \leq x \Big] - \Phi(x) \to^p 0.
\end{equation}
(Also see
Lemma~\ref{L2}.)

>>>>

Theorem~\ref{main-bootstrap-clt} presents the main result, asymptotic normality of
the distribution of bootstrapped sums. This paper adopts the
standard notation that $\E^{*}$, $\var^{*}$, etc.\ are the usual
operators with respect to the probability measure induced by the
bootstrap and will use explicit stochastic array notation
for precision.  Also note that all results are presented for the
scalar case but generalize immediately to random vectors.  All of the
proofs are presented in the appendix; only proofs for the Stationary
Bootstrap are presented, since proofs for the other methods are
similar and easier to construct.  All limits are taken as $n \to \infty$
unless otherwise noted and $\lVert \cdot \rVert_r$ denotes the
$L_r$-norm.

\begin{thm}\label{main-bootstrap-clt}
  Suppose the following conditions hold.
  \begin{enumerate}
  \item $X_{nt}$ is $L_2$-\ned\ of size $-1/2$ on an array
    $\{V_{nt}\}$ that is either strong mixing of size $-r/(r-2)$ or
    uniform mixing of size $-r/2(r-1)$, with $r > 2$.  The
    \ned\ magnitude indices are denoted $\{d_{nt}\}$.
  \item The array $\mu_{nt} - \bar \mu_n$ is uniformly bounded
    and $\sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2 \to 0$,
    where $\E X_{nt} = \mu_{nt}$ and $\bar{\mu}_n = n^{-1} \sum_{t=1}^n
    \mu_{nt}$. Moreover, $\sqrt{n} \| \bar{X}_{n} - \bar\mu_n \|_2
    \to \sigma > 0$, with $\bar X_n = (1/n) \sum_{t=1}^n X_{nt}$.
  \item There exists an array of positive real numbers $\{c_{nt}\}$
    such that $(X_{nt} - \mu_{nt})/c_{nt}$ is uniformly $L_r$-bounded
    and $c_{nt}$ and $d_{nt}/c_{nt}$ are uniformly bounded in $n$
    and~$t$.
  \item $X_{nt}^{*}$ is generated by the Stationary Bootstrap with
    geometric block lengths with success probability $p_n$, $p_n = c
    n^{-a}$ and $a,c \in (0,1)$, or by the Moving or Circular Block
    bootstrap with block length $M_n$ such that $M_n \sim n^a$ for
    $a \in (0,1)$.  Let $M_{ni}$ be the block length of the $i$th
    block, $i=1,\dots,J_n$, and define $K_{n0} = 0$ and $K_{nj} =
    \sum_{i=1}^j M_{ni}$. The last block, $M_{n,J_n}$, is defined as
    $K_{n,J_n} - K_{n,J_n-1}$, so $K_{n,J_n} = n$ a.s.
  \end{enumerate}
  Then $\hat\sigma^{*2} \to^p \sigma^2$,
  \begin{equation}\label{eq:36}
    \sup_x \big\lvert \pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \E^* \bar X_n^*) \leq x \big]
    - \pr \big[\sqrt{n}(\bar X_{n} - \E \bar X_n) \leq x \big] \big\rvert \to^p 0,
  \end{equation}
  and
  \begin{equation}
    \label{eq:2}
    \sup_x \big\lvert \pr^{*}\big[
    \sqrt{n}(\bar X_{n}^{*} -  \E^* \bar X_n^*) \big/ \hat\sigma_n^{*}
    \leq x \big] - \Phi(x) \big\rvert \to^p 0,
  \end{equation}
  where $\Phi$ is the \allcaps{cdf} of the Standard Normal
  distribution and
    \begin{equation}
      \label{eq:3}
      \hat{\sigma}_n^{*2} = \tfrac{1}{n} \sum_{j=1}^{J_n}
      \Big\{\sum_{t=K_{n,j-1}+1}^{K_{nj}} (X_{nt}^{*} - \bar X_{n}^{*})\Big\}^2.
    \end{equation}
\end{thm}

@@

Theorem~\ref{main-bootstrap-clt} relies on a general insight about the
variance of the sample mean under the bootstrap-induced
distribution. It is well-known that a key step in proving the \clt\
for arbitrary dependent processes is demonstrating that the squared
elements converge to a positive and finite limit; i.e.\ if
$\{Z_{nj};~j=1,\dots,J_n\}$ is a representative stochastic array,
$\sum_j Z_{nj}^2 \to^p \sigma^2$ is an important necessary condition
for $\sum_j Z_{nj} \to^d N(0,\sigma^2)$.  (See Section 3.2 of
\citealp{HaH:80}, for further discussion.)  For martingale difference
arrays, each $Z_{nj}$ is one of the original random variables $X_{nt}$
(typically normalized by $1/\sqrt{n}$), but for other forms of
dependence (\ned\ or mixingale arrays, for example, as in
\citealp{Jon:97}) each $Z_{nj}$ is a contiguous block of the original
random variables,
\begin{equation*}
  Z_{nj} = \tfrac{1}{\sqrt{n}} \sum_{t=(j-1) M_n +1}^{j M_n} (X_{nt} - \mu_{nt}),
\end{equation*}
that adds up to the original summation (plus potentially a negligible
residual) so $\sum_j Z_{nj}= (1/\sqrt{n}) \sum_t (X_{nt} - \mu_{nt}) +
o_p(1)$. In \citet{Jon:97}, for example, the \clt\ for mixingale
arrays assumes that there exists a sequence of blocks such that $\sum_j
Z_{nj}^2$ converges i.p., and the \ned\ \clt\ establishes conditions
under which such a $Z_{nj}$ exists.

Our insight is that the expectation of squared blocks of the bootstrap
process can be expressed as a sequence of \emph{contiguous} blocks of
the original process, so the arguments that establish convergence of
the original squared blocks can be applied with only minor changes to
the bootstrapped blocks. Consider the Moving Blocks Bootstrap,%
\footnote{To make this presentation as simple as possible, assume for
  now that $n = M_n J_n$ exactly.} %
for example, and let
\begin{equation*}
  Z_{nj}^* = \tfrac{1}{\sqrt{n}} \sum_{t=(j-1) M_n + 1}^{j M_n} (X_{nt}^* - \bar X_n).
\end{equation*}
Then, conditional on the data, the $Z_{nj}^{*2}$ are independent can
be expected to obey an \lln, so %
$\sum_j (Z_{nj}^{*2} - \E^* Z_{nj}^{*2}) \to^p 0$ %
and the CLT for the bootstrapped array requires $\sum_j \E^*
Z_{nj}^{*2}$ to converge to a positive and finite limit.  But, since
$\E^*$ only averages over the starting point of each block, we have
\begin{align*}
  \E^* Z_{nj}^{*2} &=
  \tfrac{1}{n} \sum_{\tau = 0}^{n-M_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= \tau + 1}^{\tau + M_n} (X_{nt} - \bar X_n) \Big)^2 \\
  &= \tfrac{1}{n} \sum_{\tau_0 = 0}^{M_n-1}
  \sum_{j=1}^{J_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= (j - 1) M_n + \tau_0 + 1}^{j M_n + \tau_0} (X_{nt} - \bar X_n) \Big)^2
\end{align*}
after grouping blocks separated by $M_n$ periods. For each $\tau_0$,
the summation can be expected to converge in probability through the
same arguments that were used to establish the \clt\ for the original
array.%
\footnote{Some of the details of the argument will typically need to
  change because the original \clt\ only requires convergence for
  $\tau_0 = 0$, but these details are often incidental to the original
  argument.}%
\footnote{Lemmas~\ref{bootstrap-variance-convergence-lemma} and
  \ref{bootstrapped-zsq-uniform-integrability-lemma} are particularly
  strong demonstrations of this argument.} %
A similar representation is available for the Circular and Stationary
bootstraps.

In short, the basic approach that we use to prove Theorem~\ref{main-bootstrap-clt}
is based on a fundamental connection between the second moments of the
bootstrap process and the sum of squared blocks of the original
array. Even though the details of our proof rely on specific
techniques for \ned\ arrays, this connection implies that block
bootstraps are typically consistent when the original dependent array
obeys the \clt\ and the connection should be useful for proving
consistency of the bootstrap under other dependence conditions.

Theorem~\ref{main-bootstrap-clt} can also be extended to give an \fclt\
using arguments from \citet{JoD:00b}. We show in
Theorem~\ref{main-bootstrap-invariance-theorem} that the partial sum
of the bootstrapped process obeys an \fclt\ and can be used to derive
critical values for other test statistics under the same assumptions
as Theorem~\ref{main-bootstrap-clt}. For this result, define the
following partial sums,
\begin{align*}
W_n(\gamma) &=
\tfrac{1}{\sqrt{n}} \sum_{t=1}^{\lfloor \gamma n \rfloor}
(X_{nt} - \bar\mu_n) &&\text{and}&
W_n^{*}(\gamma) &= \tfrac{1}{\sqrt{n}}
\sum_{t=1}^{\lfloor \gamma n \rfloor}
(X^*_{nt} - \mu_n^{*}).
\end{align*}
Also let $W$ denote standard Brownian Motion and $\sigma W$ denote
Brownian Motion scaled by the constant $\sigma$.

\begin{thm}\label{main-bootstrap-invariance-theorem}
Suppose that the conditions of Theorem~\ref{main-bootstrap-clt} hold
and let $\dist$ be any distance function that metricizes weak
convergence. Then
\begin{equation}\label{eq:7}
\pr^{*}[\dist(W_n^{*}, \sigma W) > \delta] \to^p 0
\end{equation}
for any positive  $\delta$.
If, in addition, $\sup_{t=1,\dots,n}|\mu_{nt} -
\bar \mu_n| = o(1/\sqrt{n})$ and
\begin{equation}\label{eq:4}
n^{-1} \sum_{s,t=1}^{\lfloor \gamma n \rfloor} \cov(X_{ns}, X_{nt}) \to \sigma^2 \gamma
\end{equation}
for all $\gamma \in [0,1]$, then
\begin{equation}\label{eq:8}
\pr[\dist(W_n, \sigma W) > \delta] \to 0
\end{equation}
for any positive $\delta$.
\end{thm}
If both~\eqref{eq:7} and~\eqref{eq:8} hold then the bootstrap can be
used to approximate the distribution of partial sums.
Note that Theorem~\ref{main-bootstrap-invariance-theorem} imposes
stronger assumptions for the original process
than for the bootstrapped process. Without~\eqref{eq:4}, the
partial sum of the original observations converges to a transformed
Brownian Motion with a different covariance process. The bootstrapped
partial sum, on the other hand, always converges to standard (but
potentially scaled) Brownian Motion because the resampling strategies
ensure that the bootstrapped process is globally covariance stationary.

@@

\subsection*{Proof of Theorem~\ref{main-bootstrap-clt}}
{%
\newcommand{\Db}[1][j]{D_{n#1}^{*}}
\newcommand{\Zsum}{\sum_{j=1}^{J_n} \Zb / \sigma_n}
First we prove that
\begin{equation}\label{eq:37}
  \sup_x \big\lvert \pr^{*}\big[
  \sqrt{n}(\bar X_{n}^{*} -  \bar X_n) \big/ \sigma_n^{*}
  \leq x \big] - \Phi(x) \big\rvert \to^p 0
\end{equation}
where $\sigma_n^{*2} = n \E^* (\bar X_n^* - \bar X_n)^2$.
Rewrite $\sqrt{n}(\bar X_{n}^{*} - \bar X_n)$ as in
Equation~\eqref{eq:16}, so
\begin{equation*}
  \tfrac{1}{\sqrt{n}} \sum_{t=1}^n (X_{nt}^* - \bar X_n) =
  \sum_{j=1}^{J_n} Z_{nj}^*
\end{equation*}
and $\{\Zb / \sigma_n^*, \Gs_{nj}\}$ is a martingale difference
array. Moreover,
\begin{equation}
  \label{eq:18}
  \Pm^{*} \Big[\sum_{j=1}^{J_n} \Zb / \sigma_n^* \leq x \Big] - \Phi(x) \to^p 0,
\end{equation}
for all $x$ if $\sigma_n^{*2} \to^p \sigma^2$ (which ensures that
$\sigma_n^{*2}$ is uniformly a.s.\ positive and holds by
Lemma~\ref{bootstrap-variance-convergence-lemma}) and the following
two conditions hold for all positive $\epsilon$:
\begin{equation}
  \label{eq:19}
  \sum_{j=1}^{J_n} \Em^{*} \big(Z_{nj}^{*2} \ind\{Z_{nj}^{*2}  >
  \epsilon\}\big) \to^p 0
\end{equation}
and
\begin{equation}
  \label{eq:20}
  \Pm^{*}\Big[\ \Big|\sum_{j=1}^{J_n} Z_{nj}^{*2} - \sigma_n^{*2}
  \Big|\ > \epsilon \Big] \to^p 0
\end{equation}
since~\eqref{eq:19} and~\eqref{eq:20} ensure that $\Zb/\sigma_n^*$ obeys
a martingale difference \clt\ \citep[e.g.][Theorem 3.3]{HaH:80}.%
\footnote{Conditional on $X_{n1},\dots,X_{nn}$, $J_n$, and
  $M_{n1},\dots,M_{n,J_n}$, the only stochastic components of $\Zsum$
  are the start periods of each block, which are
  $\discreteuniform(1,\dots,n)$ and are independent of all of the
  other random variables in the information set used for conditioning.
  Consequently, $\Pm^*$ is a regular conditional probability and
  arguments like Hall and Heyde's (1980) Theorem 3.3
  apply without modification on this probability measure. See
  also Section 23.2 of \citet{Vaa:00}.
  Also note that Hall and Heyde's Theorem 3.3 as stated imposes an
  additional restriction on the sigma-fields. However, as Hall and
  Heyde discuss on pages 59 and 63--64, that condition is unnecessary
  here because $\sigma_n^{*2}$ is measurable with respect to
  all of the $\Gs_{nj}$.} %

For~\eqref{eq:20}, $Z_{nj}^*$ and $Z_{nk}^*$ (when $k \neq j$) are
conditionally uncorrelated given $X_{n1},\dots,X_{nn},$ and $\Ms$, which
implies
\begin{equation*}
  \sum_{j=1}^{J_n} Z_{nj}^{*2} - \sigma_n^{*2} =
  \sum_{j=1}^{J_n} \Big( Z_{nj}^{*2} - (1/J_n) \E^* \sum_{j=1}^{J_n} Z_{nj}^{*2} \Big)
\end{equation*}
almost surely. But
\begin{equation*}
   \Big\{\tfrac{n}{M_{nj}} \Big(Z_{nj}^{*2} - (1/J_n) \E^* \sum_{j=1}^{J_n} Z_{nj}^{*2}\Big),
   \ \Gs_{nj} \Big\}
\end{equation*}
is a uniformly-integrable martingale difference array
by Lemma~\ref{bootstrapped-zsq-uniform-integrability-lemma}
and satisfies the \lln\ (e.g., \citealp{Dav:94}, Theorem 19.7). So
this sum converges to zero in conditional probability, proving~\eqref{eq:20}.

To prove~\eqref{eq:19}, it suffices to show that
\begin{equation}\label{eq:21}
  \Em \sum_{j=1}^{J_n} \Em^{*}\big(Z_{nj}^{*2} \ind\{Z_{nj}^{*2} > \epsilon\}\big) \to^p 0.
\end{equation}
For any $j$ and $m$ and for large enough $n$,
\begin{align}
\Em \Em^{*}\big((Z_n(u_{nj}, m)^2n/m) &\ind\{Z_n(u_{nj}, m)^2n/m > \epsilon n/m\}\big) \notag \\
&= \E \big((Z_n^{*}(0, m)^2n/m) \ind\{Z_n^{*}(0, m)^2n/m > \epsilon n/m\}\big) \notag \\
&\leq B(\epsilon n/m)\label{eq:5}
\end{align}
where $B$ is some finite monotone function such that $B(x) \to 0$ as
$x \to \infty$; the existence of a function $B$ that satisfies these
conditions (and does not depend on $n$ or $m$) is a consequence of
Lemma~\ref{bootstrapped-zsq-uniform-integrability-lemma},
Equation~\eqref{eq:30}. Since $J_n$ and $M_{nj}$ are both $\Ms$-measureable
and independent of the $u_{nj}$'s and $X_{nt}$'s,~\eqref{eq:5} implies that
\begin{align*}
\Em &\sum_{j=1}^{J_n} \Em^{*}\big(Z_{nj}^{*2} \ind\{Z_{nj}^{*2} > \epsilon\}\big) \\
& = \sum_{j=1}^{J_n} (M_{nj}/n) \Em \Em^{*}
\big((Z_n(u_{nj}, M_{nj})^2 n/M_{nj})
\ind\{Z_n(u_{nj}, M_{nj})^2n/M_{nj} > \epsilon n/M_{nj}\}\big) \\
& \leq \sum_{j=1}^{J_n} (M_{nj}/n) \, B(\epsilon n/M_{nj}) \\
& \leq \max_{j = 1,\dots,J_n} B(\epsilon n / M_{nj}) \sum_{i=1}^{J_n} M_{ni} / n \\
& = B(\epsilon n / \max_{j = 1,\dots,J_n} M_{nj}) \\
& \to^p 0 \text{\ as\ } n \to \infty.
\end{align*}
The first equality follows from simple algerbra and the definition
of the $u_{nj}$'s. The first inequality is a
consequence of~\eqref{eq:5}.  The second equality holds by
monotonicity of $B$. And convergence in probability holds by
Lemma~\ref{block-length-convergence-lemma}, completing the proof
of~\eqref{eq:19}. The Dominated Convergence Theorem and~(\ref{eq:18})
then ensure that
\begin{equation}\label{eq:22}
  \pr^{*} \Big[\sum_{j=1}^{J_n}
  Z_{nj}^{*} / \sigma_n^* \leq x \Big] - \Phi(x) \to^p 0.
\end{equation}
(Also see
Lemma~\ref{equivalent-convergence-from-lie-lemma}.)
